$\pi$ with leftovers: \\ a mechanisation in Agda

### 2020-05-03 paper/paper.tex
Go over contributions

\subsection{Contributions}
Our contributions are the following:
\begin{itemize}
\item \textbf{Formalisation of the syntax and the semantics of the \picalc{}}:
  \autoref{syntax} uses type level de Bruijn indices \cite{deBruijn1972} to introduce a syntax that is well-scoped by construction: every free variable is accounted for in the type of the process that uses it.
  \autoref{semantics} provides an operational semantics for the untyped \picalc{}, prior to any typing.
  This operational semantics is given as a reduction relation on processes.
  The reduction relation tracks at the type level what channel the communication occurs on, so that this information can then be used to state subject reduction --- aka type preservation.
  It is also defined modulo structural congruence --- a relation that acts as a quotient type that gets rid of the unnecessary syntactic minutiae introduced by the syntax of the \picalc{}.

  \item \textbf{Leftover typing for the \picalc{}, with linear, gradual and shared types}:
  \autoref{type-system} provides the type system for a resource-aware \picalc{}.
  The type system is \emph{resource-aware} in that it allows the user to provide a resource-aware algebra that is then applied to the type-system.
  Linear types, gradual types and shared types are all instances of such a resource-aware algebra too.
  Any \emph{partial commutative monoid} that is cancellative and has a minimal element is a valid such algebra.
  Multiple of these algebras can be applied to a single type system --- contexts keep information about what algebra should be applied to which element.

@@

  \item \textbf{Machine verified proofs of subject reduction and auxiliary theorems}:
  \autoref{type-safety} builds on several theorems to proof subject reduction for our type system.
  Here too, leftover typing renders the proofs for some of these lemmas straightforward, while for others it makes it possible to state them in the first place.
  We prove \textbf{framing} (well-typedness is independent of the unused resources of a process), \textbf{weakening} (new variables that are added to both the input and output usage context preserve well-typedness), and \textbf{strengthening} (removing the variables that are unused by a process preserves well-typedness).
  We use these to prove \textbf{subject congruence} (structural congruence rules can be freely applied while preserving well-typedness) and \textbf{substitution} (a variable can be substituted for another as long as this other variable is capable of doing everything the original one did), and use these to prove \textbf{subject reduction} (a well-typed process that reduces will still be well-typed).

>>>>

\subsection{Contributions}
Our contributions are the following:
\begin{itemize}
\item \textbf{Formalisation of the syntax and the semantics of the \picalc{}}:
  \autoref{syntax} uses type level de Bruijn indices \cite{deBruijn1972} to introduce a syntax that is well-scoped by construction: every free variable is accounted for in the type of the process that uses it.
  \autoref{semantics} provides an operational semantics for the \picalc{}, prior to any typing.
  This operational semantics is given as a reduction relation on processes.
  The reduction relation tracks at the type level what channel the communication occurs on, so that this information can then be used to state subject reduction --- aka type preservation.
  The relation is defined modulo structural congruence --- a relation that acts as a quotient type that gets rid of the unnecessary syntactic minutiae introduced by the syntax of the \picalc{}.

  \item \textbf{Leftover typing for the \picalc{}, with linear, gradual and shared types}:
  \autoref{type-system} provides the type system for a resource-aware \picalc{}.
  The type system is \emph{resource-aware} in that it allows the user to provide a resource-aware algebra that is then applied to the type-system.
  Linear types, gradual types and shared types are all instances of such a resource-aware algebra.
  Any \emph{partial commutative monoid} that is \emph{functional}, \emph{cancellative} and has a minimal element is a valid such algebra.
  Multiple algebras can be simultaneously used in a single type system --- usage contexts keep information about what algebra to use on which element.

@@

  \item \textbf{Machine verified proofs of subject reduction and auxiliary theorems}:
  Leftover typing allows \emph{framing}, \emph{weakening} and \emph{strengthening} theorems to be stated.
  We use these to prove \emph{subject congruence}, and together with \emph{substitution}, to prove \emph{subject reduction}.
